Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-06 popl24/composition.tex
checkpoint

When we consider internal components and hence intensional properties
of circuits, we will consider behaviors with respect to particular bits
of randomness. Hence we will consider basic program distributions. 
\begin{definition}
  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
  $\prog$ as defined in Definition \ref{def-progd}. For any program $\prog =
  (\eassign{v_1}{\be_1};\ldots;\eassign{v_n}{\be_n})$, define
  $\vdefs(\prog) \defeq \{ v_1,\ldots,v_n \}$.
\end{definition}
We also need to formalize notions of probabilistic independence, aka
\emph{separation}, and correlation. We borrow the symbols $*$ and $\sim$
from \cite{barthe2019probabilistic} with the same denotations.
\cnote{I switched here to using $\pmf$ to denote an arbitrary probability mass
function, will update earlier sections with this notation.}
\begin{definition}
  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
  We write $\sep{\pmf}{Y}{Z}$ iff for all
    $\store \in \mems(X \cup Y)$:
  $$(\margd{\pmf}{Y \cup Z})(\store) =
  \pmf(\store_Y) * \pmf(\store_Z)$$ 
\end{definition}

@@

The following Lemma allows us to separate variables in programs that may not occur
in the same subprogram, and hence consider gate certification properties in various
contexts. It is reminiscent of the \emph{frame} rule of \cite{barthe2019probabilistic}.
\begin{lemma}
  \label{lemma-gmw-frame}
  If $\sep{\progtt(\prog_1;\prog_2)}{Y}{V}$ for $V \in \vdefs(\prog_2)$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{V}$. 
\end{lemma}
\begin{proof}
  By assumption and Lemma \ref{lemma-sepjoin}, for all
  $\store^1,\store^2 \in \runs(\prog_1;\prog_2)$ we have $\store^1_Y
  \cap \store^2_V \subseteq \store$ for some $\store_a \in
  \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_Y \subseteq
  \store$ for some $\store_b \in \runs(\prog_1;\prog_2)$.  Thus, since
  we assume $X \not\in \vars(\prog_2)$ we have $\store_a \cap \store_b
  \in \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_V =
  \store_a \cap \store_b$, obtaining the result by
  \ref{lemma-sepjoin}.
\end{proof}

>>>>

When we consider internal components and hence intensional properties
of circuits, we will consider behaviors with respect to particular bits
of randomness. Hence we will consider basic program distributions. 
\begin{definition}
  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
  $\prog$ as defined in Definition \ref{def-progd}. For any program $\prog =
  (\eassign{v_1}{\be_1};\ldots;\eassign{v_n}{\be_n})$, define
  $\vdefs(\prog) \defeq \{ v_1,\ldots,v_n \}$.
\end{definition}
We also need to formalize notions of probabilistic independence, aka
\emph{separation}, and correlation. We borrow the symbols $*$ and $\sim$
from \cite{barthe2019probabilistic} with the same denotations.
\begin{definition}
  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
  We write $\sep{\pmf}{Y}{Z}$ iff for all
    $\store \in \mems(X \cup Y)$:
  $$\margd{\pmf}{Y \cup Z}(\store) =
  \pmf(\store_Y) * \pmf(\store_Z)$$ 
\end{definition}

@@

The following Lemma allows us to separate variables in programs that may not occur
in the same subprogram, and hence consider gate certification properties in various
contexts. It is reminiscent of the \emph{frame} rule of \cite{barthe2019probabilistic}.
\begin{lemma}
  \label{lemma-gmw-frame}
  If $\sep{\progtt(\prog_1;\prog_2)}{Y}{V}$ for $Y = \vars(\prog_1) \cap \vars(\prog_2)$ and
  $V \in \vdefs(\prog_2)$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{V}$. 
\end{lemma}
\begin{proof}
  By assumption and Lemma \ref{lemma-sepjoin}, for all
  $\store^1,\store^2 \in \runs(\prog_1;\prog_2)$ we have $\store^1_Y
  \cap \store^2_V \subseteq \store$ for some $\store_a \in
  \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_Y \subseteq
  \store$ for some $\store_b \in \runs(\prog_1;\prog_2)$.  Thus, since
  we assume $X \not\in \vars(\prog_2)$ we have $\store_a \cap \store_b
  \in \runs(\prog_1;\prog_2)$ and $\store^1_X \cap \store^2_V =
  \store_a \cap \store_b$, obtaining the result by
  \ref{lemma-sepjoin}.
\end{proof}
